---
title: "Finding errors in R"
date: "2018-10-06T00:00:00+00:00"
tags: ["R"]
categories: ["debugging"]
author: ["Bruna Wundervald"]
draft: false
share: false
profile: false
summary: "Summary of my post."
editor_options: 
  chunk_output_type: console
---

*Portuguese version of this same post [here](https://www.curso-r.com/blog/2018-09-03-funcoes/)!

On a daily basis, I come across people on Facebook groups, Telegram
or Twitter having troubles with some tasks in R. These
problems have two main origins:

  1. difficulties in constructing the algorithm to reach the desired result.
  2. difficulties in understanding the behavior of some function. 

The focus of this post is to give some notion about how to 
solve the second one: problems using functions. To give a little
bit of context, we can say that R packages are made by 
the community, voluntarily. Therefore, not all package documentations 
are written in the clearest way possible.  Also, currently, 
CRAN only demands a certain amount of rigor regarding the 
functions of a package, and not its documentation. This 
situation is changing for better, which can be noted 
by the existence of recommendations such as the following, 
extracted from [Hadley's book about R packages](http://r-pkgs.had.co.nz/):


> Documentation is one of the most important aspects of a good package. Without it, users won’t know how to use your package. Documentation is also useful for future-you (so you remember what your functions were supposed to do), and for developers extending your package.

While documentation is not perfect, users need to try to understand
the unexpected errors using other resources. The method I'll 
describe now is basically looking into the source code of 
functions and search for what is generating the troubles. 

There are functions that can be directly viewed in `console`, 
by just printing them without the final `()`, for example: 

```{r}
mean <- function(x, y){
  n <- 2
  sum_vars <- x + y
  mean <- sum_vars/n
  mean
}

mean
```

Just by running the name of the function we access
the code that composes it. Then if we try, 
for example:

```{r, error=TRUE}
mean(2, "1")
```


Well, what does that mean? For the new user, it might not be 
so obvious, but sum 2 and "1" doesn't work in R. Although they simply look
like two numbers, they're actually two different objects, with 
two different classes - 2 is a numeric, but "1" is a character. 

The strategy here is just to copy the function code 
and run line by line until the error is found. Remember to 
create the necessary objects first (the arguments the function
receives), like this: 

```{r, error = TRUE}
x <- 2              # The first argument of the function
y <- "1"            # The second argument of the function

# Code of the created 'mean' function   
n <- 2
sum_vars <- x + y  # error found here!
mean <- sum_vars/n
```

This was pretty simple because the function has just a few lines of code, 
and we get the error already in the second line. Now, 
is way easier to think about the probable reason of the problem
than just trying to interpret the error message. Sometimes, the 
question is exactly about using objects with the wrong structure,
and this is also what causes the weirdest errors.

The example above is with just a simple, short function. Usually, 
you'll come across extensive functions or that when printed in the
`console` won't show the code, but the `method` used, for
example: 

```{r}
# The mean function that exists in R base
base::mean
```

And how we interpret that? We're actually looking at a generic
function of the S3 class, that calls a method to use. There are 
different methods for different object types. 

Let's use  `base::mean` as an example. Which are the objects
that can be used in this function? In general, we use 
numeric vectors, but the function can also deal with other types, 
such as dates. What we mean with this is that the **same function 
can perform the same task with different object types**, by 
calling the specific **methods**. 

And what are the methods? A method is a function associated 
with a particular type of object. We can check the available methods with:

```{r, warning=FALSE}
methods(mean)
```

(depending on the package, the methods aren't exported. If you're 
dealing with one of those, try using the operator `:::`, as 
`dplyr:::filter.tbl_df`, for example!)

Now we know which methods can be used for the function 
`base::mean`. We can see the code by running the 
combination `desired-function.specific-method`, without 
the final `()`: 

```{r}
base::mean.default
```

Let's get back to the discovery of error sources. Say we run the
following line of code, which won't go so well: 


```{r, error=TRUE}
base::mean(c("1", 3))
```

Note that this is not an error, but a `warning` message. Anyway, 
the `NA` is certainly not the desired output when trying to obtain
the mean of two numbers. What happened? We'll check it by
using the code of  `base::mean.default`: 


```{r, eval = TRUE}
# Defining the vector which is the argument of the function
x <- c("1", 3)

# Code of  base::mean.default 
if (!is.numeric(x) && !is.complex(x) && !is.logical(x)) {
  
  warning("argument is not numeric or logical: returning NA")
  return(NA_real_)
}

# We can stop here as the error was already found!
```

In the first line of the function, we have the indication 
of the problem: the vector is neither numeric, complex or logical. If
we want the function to work, is must receive those types of objects.
Well, now is easier to understand what was done wrong! 

# Why should I try to understand the errors?

1. Making mistakes is a great way to learn. 

Broadly, in my experience with searching errors in functions, I 
always end up learning something new. The own correction of the error
generally leads me to know something that wasn't quite 
obvious before. The errors make me look for a better understanding 
of their existence, which hence leads to an improvement in 
the comprehension about programming logic and R in general. 

2. Your own independence. 

Is always more efficient being able to solve your own problems. Not
rarely, I see questions about R in the internet
that take a long time to have a proper answer. With a little deeper 
search for the source of your errors and its consequent solution, 
this waiting will be avoided (not that you shouldn't make 
questions, of course). 

3. Exposition to the diversity of "coding styles" we have in 
the community. 

Particularly, I can comment that, since R packages are made by the
community, there is a wide diversity in the ways they're written. 
The contact with this diversity, aka going deep in how functions
are coded, leads me not only to learn more about R but also about
how to refine my own coding style. 

# Wrap-up

In this blog post, I explained a bit about how to search for errors
in R functions. We talked about:

  - how to see the source code of simple functions; 
  - how to see the source code of the S3 class functions;
  - how to use these code for indentifying the erros;
  - how we can learn with our own mistakes; 
  - how we can save time by learning about finding & fixing errors. 

That's it for now. I hope you liked it. Any questions can be addressed
to me directly or here via comments! 
